\TITLE

### 2021-01-27 draft/response-3.tex

\makecell{Future\\Case} &
\makecell{Past\\Death} &
\makecell{Now\\Death} &
\makecell{Future\\Death} \\
\midrule
PVI                     & -0.425      & -0.292      & -0.124      & -0.434      & -0.373      & -0.239    \\
INSURANCE               & 0.018       & 0.017       & 0.012       & 0.018       & -0.002      & 0.001     \\
NBEDS                   & 0.136       & 0.131       & 0.217       & 0.089       & 0.080        & 0.109     \\
NBEDS/1000              & 0.138       & 0.121       & 0.209       & 0.092       & 0.069       & 0.088     \\
NHOSP                   & 0.074       & 0.119       & 0.165       & -0.009      & 0.028       & 0.021     \\
ESTBEDS                 & 0.136       & 0.131       & 0.217       & 0.089       & 0.080        & 0.109     \\
SENIOR                  & 0.104       & 0.028       & 0.033       & 0.152       & 0.164       & 0.156     \\
BLACK                   & 0.209       & 0.099       & 0.208       & 0.216       & 0.085       & 0.116     \\
HISPANIC                & -0.072      & -0.032      & -0.064      & -0.091      & -0.068      & -0.046    \\
POP\_DENSITY\_2010      & -0.302      & -0.210      & -0.175      & -0.272      & -0.231      & -0.216    \\
POVERTY                 & 0.128       & 0.059       & 0.200       & 0.148       & 0.031       & 0.080     \\
SVI\_MINORITY           & 0.029       & 0.069       & 0.081       & 0.010        & 0.006       & 0.087     \\
SVI\_OVERALL            & 0.140       & 0.158       & 0.238       & 0.125       & 0.113       & 0.179     \\
CASTHMA                 & 0.238       & 0.271       & 0.307       & 0.205       & 0.288       & 0.309     \\
HIGHCHOL                & 0.202       & 0.185       & 0.131       & 0.192       & 0.243       & 0.224     \\
DIABETES                & 0.200       & 0.200     & 0.178         & 0.206       & 0.232       & 0.230     \\
OBESITY                 & 0.199       & 0.205       & 0.194       & 0.191       & 0.229       & 0.212     \\
CANCER                  & -0.145      & -0.102      & -0.129      & -0.138      & -0.058      & -0.059    \\
STROKE                  & 0.123       & 0.125       & 0.117       & 0.167       & 0.143       & 0.135     \\
MHLTH                   & 0.110       & 0.161       & 0.186       & 0.076       & 0.175       & 0.222     \\
CSMOKING                & 0.189       & 0.258       & 0.249       & 0.160        & 0.289       & 0.323     \\
CHOLSCREEN              & 0.164       & 0.001       & -0.026      & 0.187       & 0.070       & 0.038     \\
INSURANCE               & 0.018       & 0.017       & 0.012       & 0.018       & -0.002      & 0.001     \\
CHD                     & 0.225       & 0.284       & 0.233       & 0.178       & 0.327       & 0.337     \\
CHECKUP                 & 0.206       & 0.221       & 0.280       & 0.201       & 0.272       & 0.343     \\
KIDNEY                  & 0.045       & 0.028       & 0.032       & 0.022       & 0.071       & 0.053     \\
BINGE                   & 0.031       & 0.044       & 0.009       & 0.083       & 0.030       & 0.022     \\
LPA                     & 0.191       & 0.187       & 0.197       & 0.147       & 0.220       & 0.214     \\
ARTHRITIS               & 0.046       & 0.013       & -0.021      & 0.061       & 0.089       & 0.079     \\
BPMED                   & 0.060       & 0.080        & 0.124       & 0.056       & 0.073      & 0.090      \\
PHLTH                   & 0.064       & 0.132       & 0.104       & 0.022       & 0.129       & 0.114     \\
BPHIGH                  & 0.056       & 0.051       & 0.048       & 0.058       & 0.084       & 0.068     \\
COPD                    & 0.163       & 0.163       & 0.153       & 0.175       & 0.220       & 0.240     \\
NORM\_POP               & -0.305      & -0.258      & -0.218      & -0.193      & -0.218      & -0.099  \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\end{comment}

\section{Conclusion}

While this study focuses on if and which risk factors can improve the
predictions, we do not focus on why. Such analysis could provide
insights into causal relationships. For example, information such as
health insurance and the number of hospitals and beds are often
related to the quality of services that affect early detection and
treatment. Furthermore, population parameters such as population density can
determine the rate of spread and the number of cases. Similarly,
distributions of comorbidities in a given population such as physical
health, cholesterol screening, diabetes, etc., are found to be
insightful in predicting outcomes \citep{Maleki}. Identifying such risk
factors and correlating them to the prediction accuracy provides
valuable candidates for future studies.

A variety of mathematical and computational models have been used for
predicting COVID-19 outcomes \citep{Jewell}. It includes compartmental
models such as the SIR (Susceptible-Infected-Recovered) and SEIR
(Susceptible-Exposed-Infected-Recovered) which are relatively easy to
compute, and thus, commonly used in epidemiology. However, their
simplistic assumptions, such as homogeneous mixing within populations,
might make them less realistic for pandemics such as COVID-19. For
instance, such models often assume a closed population that does not
change. However, COVID-19 dynamics have prominently included massive
migration, deaths due to comorbidities temporarily conferred immunity,
and notably, major disparities in terms of socio-economic determinants
of health.

In terms of implementation, SIR and SEIR models are fitted with point
estimates, which may not adequately reflect the disease dynamics. For
instance, it is difficult to account for behavioral changes (e.g.,
human mobility) or systemic inadequacies (e.g., availability of
hospital beds) that may appear in specific subpopulations. Finally,
the compartmental model estimates do not allow for quantification of
uncertainty, which makes it harder to use them in policy-making. A
comparison of modeling approaches concluded that while SEIR model
performed better for a couple of states in the U.S., alternatives have
performed better for the other states \citep{Bertozzi}. AICov has
addressed most of the above modeling issues in its design, as per the
requirements of its forecasting objectives.

We present AICov as an integrative deep learning framework for
COVID-19 forecasting with the help of population covariates. Thus, its
objectives are different from those of traditional compartmental
models. One of the important features of the architecture for AICov is
that it is by design targeting Cloud and High Performance Computing
(HPC) resources to conduct parameter sweeps to leverage sophisticated
deep learning toolkits. The architecture allows the integration of
various data sources that can update the data from its sources on
demand and update its model predictions based on newly introduced
data. Parameters can easily be adjusted via Jupyter notebooks that, in
turn, call the computational backends on HPC and cloud resources.

In addition to this architecture, we used data collected from multiple
public sources and agencies, and integrated the same across spatially
contiguous units such as cities or metropolitan areas. In our
analysis, we have focused on 110 selected cities of the U.S., but the
framework is general and can be used to analyze similar data on
pandemics from anywhere in the world.

The comprehensive analysis showcases the feasibility of our
approach. Based on the outcomes reported it can lead to an improved
prediction once we integrate risk factors in addition to the
time-dependent data such as cases and death resulting from
COVID-19. While the improvements observed were modest, they do present
a concrete way of improving the forecast models. Inclusion of further
putative factors from the ongoing worldwide studies on Covid-19 will
only strengthen the future applications of our integrative framework.

We have shown that deep learning can return very good results while
using smooth data, which we used in our empirical fits. For real data
as presented to us for the daily changes, it still produced good
results and even was able to predict changes based on weekly
fluctuations. This is typically not achieved by other non-data based
model approaches. Also, we have used in our data only cases and
deaths, but we intend to expand this by using data about recovery and,
in the future, immunization.

We have experimented with different hyperparameters and included in
this study a selection of hyperparameters that have worked well for
this data set.

As we have set up the first version of our AICov software, we have
used cloud and high-performance computers. All of our sophisticated
analyses were run in a day on at most 16 computers. All deep learning
algorithms that were run on GPUs were run on NVIDIA Tesla
K80s. However, our framework is on purpose generalized so that other
compute resources can be integrated and leveraged. This includes
local, HPC, and cloud computing resources, as well as different
GPUs. Naturally, we also need to take into consideration not only the
compute time but also the time it takes for the researcher to derive
the models. As our approach is self-learning based on automated data
updates, the time to prepare an updated model can be automatized and
minimal input is needed. Thus the overall effort is very competitive.

\begin{comment}
Furthermore, we are currently conducting experiments with multiple
risk factors at the same time to identify the most significant
combinations of them.  As we have shown that the bivariate analysis
has a locality dependency, we will explore categorizing locations with
similar demographics and conduct further experiments on demographical
similar regions.
\end{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%
% SUPLEMENTAL
%

\section*{Supplementary Materials}

The code and paper document represented to implement AICov are contained in several repositories:

\begin{enumerate}
    \item The entire cloudmesh code on which the cloud based implementation of the AICov framework is based and contains over 70 contributors is available publicly at \url{https://github.com/cloudmesh}. Cloudmesh contains a number of modules that dependent on the users access to cloud resources can be customized. A detailed manual about the configuration is available at \url{https://cloudmesh.github.io/cloudmesh-manual/}.

    \item The entire COVID-19 analysis leverages cloudmesh and uses Jupyter notebooks to coordinate its workflow as discussed in the architecture Figure \ref{fig:arch}. The code and data for the results presented in this paper are located in the repository at \url{https://github.com/cloudmesh/cloudmesh-covid}.

    The data was analysed on a variety of supercomputing resources including an allocation of 20 compute nodes that were utilized to execute the  repeated model creation to assure reproducible results.

    However, as a publication is pending, at this time, the use of the data is copyrighted and can not be used for other publications without contacting the authors. The data gathering and analysis is a significant intellectual contribution and we like to avoid that the data is taken before we have not secured a publication.

    \item The entire paper is located in \LaTeX source in the github repository \url{https://github.com/cyberaide/paper-covid}. This repository will be open sourced after acceptance of publication to not violate any publisher restrictions. If desired the authors can grant access to this repository prior to publication. Please contact the corresponding author.

\end{enumerate}

%
% ACKNOWLEDGMENT
%

\section*{Acknowledgements}

This work is partially supported by the National Science Foundation
(NSF) through awards Cyberinfrastructure Framework for 21st Century
Data Infrastructure Building Blocks (1443054), Network for
Computational Nanotechnology Engineered nanoBIO Node (1720625),
Cybertraining (1829704), CyberInfrastructure for Network Engineering
and Science (1835598) and Global Pervasive Computational Epidemiology
(1918626).  We thank J. Kadupitiya for several useful discussions on
deep learning frameworks.

\clearpage

\section*{Acronyms}

This section contains the list of acronyms in Table \ref{tab:acronyms}
as used in the paper for easy reference. The definitions of the risk
factors are however given within the paper in
Table~\ref{tab:risk-factors} on page \pageref{tab:risk-factors} and
not repeated here.

\begin{table}[!h]
  \caption{Non-risk factor acronyms used in this paper.}
  \label{tab:acronyms}
  \begin{tabular}{ll}
    \toprule
    Abbreviation & Description \\
    \midrule
    API & Application Programming Interfaces \\
    BRFSS & Behavioral Risk Factor Surveillance System \\
    CDC & U.S.~Centers for Disease Control and Prevention \\
    CI & Continious INtegration \\
    COVID-19 & COrona VIrus Disease 2019  \\
    CSSE & Center for Systems Science and Engineering \\
    cum & cumulative \\
    FIPS & Federal Information Processing Standards \\
    GPUs & Graphics Processing Units  \\
    HPC & High Performance Computing \\
    LSTM & Long Short-Term Memory \\
    MSE & Mean Squared Error \\
    NBDIF & NIST Big Data Working Group \\
    NCHS & CDC National Center for Health Statistics \\
    NIST & National Institute of Standards and Technology \\
    NONE & No risk factors used \\
    NORM\_POP & Normalized Population \\
    NSF & National Science Foundation \\
    RELU & REctified Linear Unit \\
    REST & REpresentational State Transfer  \\
    RMSE & Root Mean Squared Error \\
    RNN & Recurrent Neural Network \\
    SEIR & Susceptible, Expose, Infectious, Recovered) \\
    SELU & Scaled Exponential Linear Unit \\
    SIR &  Susceptible, Infectious, Recovered \\
    U.S. & United States of America \\
    WHO & World Health Organization \\
    \midrule
  \end{tabular}
\end{table}

\clearpage

\bibliographystyle{jdsurl}
\bibliography{bib/data-covid19,bib/lstm,bib/cloudmesh}

\end{document}

### 2021-01-27 final/response.txt

Review Response JDS2007-001RA0

Revision due: 2020-11-24

We are happy to resubmit our paper and are very thankful for the great reviews we received. Thanks to the review comments, the quality of the paper has significantly improved.

@@

In particular, we addressed the following:

*  The paper has been checked multiple times by multiple native English speakers.

*  The organization of the paper has been improved while focusing entirely on the deep learning framework. Several sections and images that distracted the reader from the paper's primary focus (which was evident by the careful reviews) were removed.

*  Reviewer 1: The method has been described using a recipe approach without giving the rationality behind it. There are too many predictive models for COVID-19. What are the advantages of the proposed method? The authors should explain their competitiveness, at least from the following two perspectives. We addressed these issues through the next two items.

*  Reviewer 1: First, compared with the very classical SIR/SEIR and their improved models, what are the advantages of the proposed model?

@@

    A new section has been added to discuss the advantages and disadvantages.

*  Reviewer 1: Is the accuracy improved significantly?

    While using the real-time data our accuracy is very good as it, for example accounts for the weekly fluctuations that are not covered by other prediction models. Moreover, the point is not only a single model in time, but that the model adapts itself over time while new data is added. This has been pointed out in the paper.

*  Reviewer 1: Second, it seems that the data used are all very structured, what about regression model framework?

@@

    The beauty of this deep learning algorithm is that it learns the trends and performs well even if the data is not smoothed with regression or a seven-day average. Instead, it even learns the weekly fluctuations which we point out in the revised version of the paper.

*  Reviewer 1: Is the performance competitive?

    The performance is very competitive and an explanation is added to the conclusion.

*  Reviewer 1: The current deep learning framework is lack of interpretation of the investigated risk factors.

@@

    We added the following section:

While this study focuses on if and which risk factors can improve the predictions, we do not focus on why. Such an analysis could provide insights into causal relationships. For example,  information such as health insurance and the number of hospitals and beds are often related to the quality of services that affect early detection and treatment. Furthermore, population parameters such as density can determine the rate of spread and the number of cases. Similarly, distributions of comorbidities in a given population such as physical health, cholesterol screening, diabetes, etc., are found to be insightful in predicting outcomes. Identifying such risk factors and correlating them to the prediction accuracy provides valuable candidates for future studies.

*  Reviewer 2: The manuscript is in a rather preliminary shape, and needs significant work before being in a publishable shape.

@@

    We believe we have addressed this issue by addressing all reviewer's comments and more.

*  Reviewer 2: Section 2.2 does not look sufficient as a subsection.  It should be either expanded to include more details, or combined with Section 2.1.3.

@@

    The sections have been improved/combined, and a section from later on in the paper was added.

* Page 5, under "Data Source Integration", what does "we include integrate data integration capabilities" mean?

@@

    The sentence has been corrected.

* Reviewer 2: 4.  Please consider writing the requirements on Pages 5 and 6 into structured paragraphs.

    The requirements have been rewritten as paragraphs. Also, we added a numbered requirement tag to them and kept the high-level name of the requirement for increased readability and structure.

* Reviewer 2: Item 5., 6., 7., 8., 9., and 10. Have been fixed by excluding the text related to them from the paper as we found they distracted too much from our main contributions.

* Reviewer 2: 12.  In Section 2.1,  the activation function is the sigmoid function.  In Section 6, RELU was used.
